---
title: "imeassure-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{imeassure-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup}
library(imeassures)
```


# Introduction
The "imeassure" package provides tools for diagnosing the influence of individual data points in a linear model. It offers three distinct methods for calculating influence measures and visualising them. These tools are designed to help users identify influential points that may affect the quality and fit of linear models, ultimately improving the robustness and accuracy of their analyses.

# Package Installation
To install the 'imeassure' package, follow these steps:
   
**1.** Before installing 'imeassure', make sure you have the 'devtools' package installed on your computer. If you don't have it installed, you can do so by running the following command in your R console:

```{r}
# install.packages("devtools")
```
**2.** Once devtools is installed, you can install imeassure directly from GitHub. In your R console, run the following command:
```{r}
# devtools::install_github("aniusk898/Assignment2DATA501")
```
This command will download and install the "imeassure" package from the specified GitHub repository.

**3.** After the installation is complete, load the package into your R session by running:

```{r}
# library(imeassure)
```
With the package loaded, you can now start using its functions. Follow the instructions provided in the sections above to begin analyzing the influence of points in your linear models.

   
   
# Functionallity

## Functions To Calculate The Measures Of Influence
To calculate the influence measures, you need to provide a linear model (lm) object as input. The function will check if the input is valid, and if there’s an issue, it will return an error message.

## "cooks_distance" Function
The "cooks_distance" function calculates Cook's distance, which is used to identify observations that have strong or significant influence on the linear model. A high Cook's distance value, indicate that the observation have a big impact on the fitted values, and these observations might need a closer look or could even be considered for removal from the model.

You can use the function in R as in the following example:
```{r}
data_valid <- data.frame(
  y = c(4.2, 5.1, 7.3, 6.5, 8.0, 6.7, 10),
  x1 = c(2.1, 3.5, 5.1, 4.3, 6.0, 5.5, 7)
)
model_valid <- lm(y ~ x1, data = data_valid)
cooks_d <- cooks_distance(model_valid)
print(cooks_d)
```
The output is a numeric vector containing the Cook's distance values for each observation in the dataset.

## "diffits" Function
The "diffits" function calculates the DFFITS value for each observation in a linear model. DFFITS helps to identify data points or observations that have a big impact on the model by measuring how much the predicted value changes when you take out an observation. A high DFFITS value, means that observation is strongly influencing the model's predictions, so you might want to take a closer look at it or even consider removing it to make the model more accurate.


You can use the function in R as in the following example:
```{r}
diffits_d <- diffits(model_valid)
print(diffits_d)
```
The output is a numeric vector containing the DFFITS values for each observation in the dataset.

## "hadis_influence" Function
The "hadis_influence" function calculates Hadi’s influence measure for each data point in a linear model. This measure helps you find observations that might be messing with how the model fits and predicts overall. If you see a high Hadi’s value, it means that data point could be having a bigger impact than you'd want, so it’s a good idea to check it out and see if it’s throwing off your analysis.

You can use the function in R as in the following example:
```{r}
hadis_influence_d<- hadis_influence(model_valid)
print(hadis_influence_d)
```
The output is a numeric vector containing the the Hadi's Influence values for each observation in the dataset.

While calculating the distances is a great starting point for identifying influential observations, a visualization can provide a clearer picture. Each influence measure has its own threshold formula to determine if an observation is highly influential. These thresholds are the same as those used in the olsrr package, and you can find more details at the following link [https://cran.r-project.org/web/packages/olsrr/vignettes/influence_measures.html](url)

## "plot_infmeasures" Function

The "plot_infmeasures" function, accepts four arguments, two of which are required and two are optional:

### Required Arguments:

**1.** model: The first argument is the linear model object (of class lm) for which you want to calculate and visualize the influence measures.

**2.** data: The second argument is the dataset used to fit the model. This should be provided as a data.frame.

### Optional Arguments:
**3.** measure: The third argument specifies which influence measure to plot. It accepts "cooks", "diffits", or "hadi". If not specified, "cooks" is used by default.

**4.**. threshold: The fourth argument allows you to set a custom threshold value for identifying high-influence points. If not provided, the function will use standard thresholds for each measure, similar to those in the olsrr package.

Before generating the plot, plot_infmeasures calls the "valInputs" function to ensure that the input data and model are valid. valInputs checks that the model is of class lm and that the data is a properly formatted data.frame. If any issues are detected, the function will return an error, ensuring that the analysis is based on correct inputs.

The plot_infmeasures function is designed to help you quickly identify influential data points through visualizations. The resulting plot shows each observation’s influence, with points exceeding the threshold highlighted, making it easier to spot potential outliers or influential cases.

Here’s an example of how to use plot_infmeasures:
```{r,fig.align='center'}
plot_infmeasures(model_valid, data_valid, measure = "cooks", threshold = NULL)
```

In this example, the function plots Cook’s distance for the model_valid, using the default threshold, which is displayed in the upper right corner of the plot. Observations that exceed this threshold (indicated by the horizontal red line) are identified as influential and are highlighted in red. These points may be considered outliers and warrant further investigation.



## Examples of Errors in the arguments
```{r}
# model_invalid <- list()
# valInputs(data_valid, model_invalid)
```

The output is "Error in `valInputs()`:
! The argument 'model' must be an object of class 'lm'."

```{r}
# model_valid <- lm(y ~ x1, data = data_valid)
# data_invalid <- as.matrix(data_valid)
# valInputs(data_invalid, model_valid)
```

The output will be is "Error in `valInputs()`:
! The data must be a data frame."

```{r,fig.align='center'}
plot_infmeasures(model_valid, data_valid, measure = "hadis_influence")
```
As in the previous measure, the observations highlited in red shows 
